\section{Curriculum Proposal}

For the purposes of cohesive programming curriculum in the Mechanical and Nucelar Engineering Department 
at KSU, proposes the use of Python in conjunction with library created for the MNE department, 

\section{Python}

Python is a multipurpose, interpreted programming langauge that emphasises code readability. The langauge is
was originally introduced by Guido van Rossum in 1991 and has gone through three major versions,
the most recent being the aptly named Python 3. Using Python, particularly for first time programmers,
comes with the following benefits:

\begin{itemize}
    \item Python is an interpreted language; therefore, users do not need to install and configure
    a compiler. The setup process for a compiler can often be confusing and a major barrier to entry
    for new users. Since Python does not have a compiler, the only download that is necessary to run 
    a .py file is Python itself. The lack of a compiler does come with some downsides, namely speed,
    but these will be addressed later on.
    \item Python is designed to be human readable. This means that the syntax of the language closely
    matches the structure of the plain English. This emphasis on readability greatly reduces the
    burden on new users and bolsters code comprehension, even in more experienced coders. Consider 
    a program that prints the phrase "Hello world!"  to the terminal. In C++, this would be done with 
    the following code:

    \begin{lstlisting}[language=C++]
    #include <iostream>

    int main() {
        std::cout << "Hello world!" << std::endl;
        return 0;
    }
    \end{lstlisting}

    Not only is this code not human readable, but it requires the use of imports, functions, return
    values, namespaces, and non-obvious syntax. In Python, however, only one line of code is needed:

    \begin{lstlisting}[language=Python]
    print("Hello world!")
    \end{lstlisting}
    \item Python is a multipurpose, general-use language. The language is adept at everything from
    data science to building a webiste to powering a microcontroller (using MicroPython). With such 
    a wide breadth of use cases, learning the language opens the door to countless different applications 
    and projects.
    \item Python is open-source and currently one of the most used languages. Thanks to its popularity,
    many tools exist to make programming with Python easier, more powerful, and more accessible such
    as integrations with Jupyter Notebooks and Visual Studio Code. However, this can potentially be 
    two edged sword, as will be discussed later.
\end{itemize}

Python provides a solid starting point for new programmers while also being fully capable of high-level,
advanced programming making it a good language for users of all skill levels. However, it does not come 
without concerns of its own, most noteably version controlling. This will be addressed at
length in a future chapter.

\section{MicroPython}

MicroPython is a slimmed version of Python that is designed to run on microcontrollers, such as the
Raspberry Pi Pico or ESP32. This Python derivative removes some high level features in exchange for
a smaller interpreter and a machine library dedicated to interacting with the hardware of the 
microcontroller, which Python would normally not interact with.

While MicroPython comes with all the benefits of Python, it also comes with one of the drawbacks
of Python: speed. Unlike traditional low level languages like C/C++, MicroPython is not compiled,
making it significantly slower than comparable languages. In some applications, the difference in
speed has no affect on the application. Others, like a control system, heavily depend on fast loop
times and are greatly impacted by speed.

\section{Raspberry Pi Pico}

The Raspberry Pi Pico is a small-but-powerful microcontroller made by Raspberry Pi. A microcontroller
is an electronic input-output device capable of running uploaded code. It is differentiated
from a microprocessor, such as a Raspberry Pi 5 or a desktop computer, by the fact that it does not run 
an operating system. 

The using a Pico, as a newer competitor in the saturated microcontroller market, begs the question of 
``why not use the Arduino Uno or ESP32?'' While all three devices support ADC, SPI, I2C, and have 
around 40 GPIO pins, only the ESP32 and Pico support MicroPython, as well as built in wireless and 
Bluetooth connectivity.

A case could be made for using the ESP32 over the Pico, given its rise in popularity in the home
automation space. However, the ESP32 comes with greater restrictions in GPIO usage and the majority
of its documentation and drivers are for C++ programs rather than MicroPython, giving the Pico an
edge in user friendlinesss.

As was the case with Python and MicroPython, the Pico does not come without drawbacks. Almost all 
sensors come with drivers written in C++ for the Arduino Uno or ESP32. This is not the case for 
the Pico. Some sensors made specifically to work with the Pico come with drivers, and others may 
be found on GitHub thanks to the work of another tinkering user, but many pieces of hardware do 
not have driver support for the Pico. This could end up being a limiting factor for students trying 
to do unguided projects.

\section{Visual Studio Code}

Visual Studio Code is a cross-platform, highly customizeable development environment created by
Microsoft. It is one of the most used and most accessible development environments available, 
and supports use with any language. VSCode can almost be thought of as a Word processor where 
spell check looks for syntax errors and keywords instead of grammatical mistakes. By installing
extensions, which can be done from inside VSCode, developers can write code for nearly any 
microcontroller or processor and in nearly any language.

For the purposes of the Mechanical Engineering Department, extensions for Python, MicroPython and
the Pico, and Jupyter Notebooks (a type of interactive programming environment that will be utilized
later) will be used. The extensions are both easy to use and install, but share a downside with
Python: version control. Since these extensions are always in active development and get updated
to work with the newest version of Python and Visual Studio Code, which may not always be desireable.
This issue will be discussed in greater detail in a later chapter.
